---
title: "DC Prelim Data"
author: "Adam R"
date: "November 3, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE
)

library(tidyverse)
library(scales)
library(modelr)

scores_all_grades <- read_csv("data/DC/dcps_parcc_scores_all_grades_17_18.csv", na = c("n < 10")) %>% 
  filter(Grade != "ALL") %>% 
  filter(`School Code` != "DCPS") %>%
  mutate(`School Code` = as.numeric(`School Code`))

scores <- read_csv("data/DC/dcps_parcc_scores_all_grades_17_18.csv", na = c("n < 10")) %>% 
  filter(Grade == "ALL") %>% #Show all grades
  select(-Grade) %>%
  filter(`School Code` != "DCPS") %>%
  mutate(`School Code` = as.numeric(`School Code`))

schools <- read_csv("data/DC/dcps_school_scorecard_17_18_trimmed.csv") 

enrollment <- read_csv("data/DC/enrollment-data-17-18.csv") %>% select(-Name)

budget <- read_csv("data/DC/dcpsbudget.ourdcschools.org-budget-data.csv") %>% 
  filter(YEAR == 2018) %>% 
  select(-WARD, SCHOOLNAME, -LAT,-LON,-YEAR ) %>%
  mutate(`Per Student Spending`=AMT_ENROLLMENT/TOTALENROLLMENT)

dc <- scores %>% left_join(schools, by=c("School Code" = "Dir_School_Code")) %>%
  left_join(enrollment, by=c("School Code" = "Code")) %>%
  mutate(ela_tests = `# of English Language Arts (ELA) Test Takers`,
         ela_proficient = `ELA # Proficient`,
         math_tests = `# Math Test Takers`,
         math_proficient = `Math - # Proficient`,
         Uniform = as.factor(Uniform)) %>%
  mutate(Safe_Truancy_Current_Year = as.numeric(sub("%", Safe_Truancy_Current_Year, replacement="")) / 100) %>%
  mutate(Perf_Teachers_Prior_Year = as.numeric(sub("%", Perf_Teachers_Prior_Year, replacement="")) / 100) %>%
  mutate(Perf_Teachers_Current_Year = as.numeric(sub("%", Perf_Teachers_Current_Year, replacement="")) / 100) %>%
  mutate(math_proficiency = math_proficient / math_tests) %>%
  mutate(Demo_Asian = as.numeric(sub("%", Demo_Asian, replacement="")) / 100) %>% 
  mutate(Demo_Black = as.numeric(sub("%", Demo_Black, replacement="")) / 100) %>% 
  mutate(Demo_Hispanic = as.numeric(sub("%", Demo_Hispanic, replacement="")) / 100) %>% 
  mutate(`Demo_N.Amer` = as.numeric(sub("%", `Demo_N.Amer`, replacement="")) / 100) %>% 
  mutate(Demo_White = as.numeric(sub("%", Demo_White, replacement="")) / 100) %>% 
  mutate(Demo_Multi = as.numeric(sub("%", Demo_Multi, replacement="")) / 100) %>% 
  mutate(Demo_NHawPI = as.numeric(sub("%", Demo_NHawPI, replacement="")) / 100) %>%
  mutate(Demo_Asian_Total = Demo_Asian * Enroll_Total) %>% 
  mutate(Demo_Black_Total = Demo_Black * Enroll_Total) %>% 
  mutate(Demo_Hispanic_Total = Demo_Hispanic * Enroll_Total) %>% 
  mutate(Demo_NAmer_Total = `Demo_N.Amer` * Enroll_Total) %>% 
  mutate(Demo_White_Total = Demo_White * Enroll_Total) %>% 
  mutate(Demo_Multi_Total = Demo_Multi * Enroll_Total) %>% 
  mutate(Demo_NHawPI_Total = Demo_NHawPI * Enroll_Total) %>%
  left_join(budget, by=c("School Code" = "SCHOOLCODE"))

```

# Notes
## Notes from August 2018 PARCC Presentation

- PARCC scores in DC up for 3rd year in a row across ELA and Math. 33.3% passing in ELA and 29.4% passing in Math.

## Notes from http://dcpsbudget.ourdcschools.org/

- While individual school budgets account for roughly 85% of total educational expenditures in the DCPS school system, they do not cover everything. Central office, textbooks, athletics, special education related services other than psychologists and social workers, utilities, security, food service, and maintenance all fall outside individual school budgets and so are not shown above.

- The result is that per pupil allocations for general education vary widely from school to school with variable correlation to enrollment. 

- The at-risk allocation is based on the number of students that are at-risk of academic failure based on one or more of the following: homeless, in the District's foster care system, receiving Temporary Assistance for Needy Families (TANF) or the Supplemental Nutrition Assistance Program (SNAP), or in high school only, are at least one year older than the expected age for their grade. According to DC law, at-risk funds are supposed to follow the student to his or her school.

## Notes from https://dcps.dc.gov/node/966292

- Total # of schools is 116 (but we only have 111 schools with PARCC results)

## Notes from Bellweather Teacher retention report
https://bellwethereducation.org/publication/retaining-high-performers-insights-dc-public-schools-teacher-exit-survey

# Summary Stats
## Enrollment
```{r}

ggplot(dc) + geom_col(width=1,aes(x=reorder(as.factor(`School Code`),`Total Enrolled`), y=`Total Enrolled`)) + coord_flip() + xlab("School Code") + ylab("Total Enrolled Students") + labs(title="Students Enrolled by School (n=111)")

```

```{r}
summary(dc$`Total Enrolled`)
```

## PARCC Results
```{r}
dc_math <- dc %>% gather(key=`Math Level`, value=`Math PARCC Outcomes`, `Math - # Level 1`, `Math - # Level 2`, `Math - # Level 3`, `Math - # Level 4`, `Math - # Level 4`, `Math - # Level 5`) %>% select(`Math Level`, `Math PARCC Outcomes`)

ggplot(dc_math) + geom_col(aes(x=`Math Level`, y=`Math PARCC Outcomes`))
```

## PARCC Results by Ward

```{r, echo=FALSE, message=FALSE, warning=FALSE}
dc %>% group_by(Dir_Ward) %>% summarise(`ELA Test Takers`=sum(`# of English Language Arts (ELA) Test Takers`, na.rm=T), 
                                             `ELA Proficient`=sum(`ELA # Proficient`, na.rm=T)) %>%
  mutate(`% ELA Proficient` = percent(`ELA Proficient` / `ELA Test Takers`)) %>%
  arrange(`% ELA Proficient`)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
dc %>% group_by(Dir_Ward) %>% summarise(`Math Test Takers`=sum(`# Math Test Takers`, na.rm=T), 
                                             `Math Proficient`=sum(`Math - # Proficient`, na.rm=T)) %>%
  mutate(`% Math Proficient` = percent(`Math Proficient` / `Math Test Takers`)) %>%
  arrange(`% Math Proficient`)
```

# Math Test Results by School Uniform

```{r}
dc %>% count(Uniform)
```

```{r}
  ggplot(dc) + geom_boxplot(aes(x = Uniform, y=  (math_proficient/math_tests)))
```

# Math Test Performance by Truancy Rate

```{r}
ggplot(dc) + geom_point(aes(x = Safe_Truancy_Current_Year, y = (math_proficient/math_tests)))
```

# Test Perfomance by Per Student Spending
## Per Student Spending by Ward
```{r}
ggplot(dc %>% select(`Per Student Spending`, Dir_Ward),aes(`Per Student Spending`, color = as.factor(Dir_Ward)), omit.na=T) + geom_freqpoly(binwidth=1000)
```

```{r}
summary(dc$`Per Student Spending`)
```

## Math

```{r}
ggplot(dc) + geom_point(aes(x=`Per Student Spending`, y=(math_proficient/math_tests)))
```

## ELA
```{r}
ggplot(dc) + geom_point(aes(x=`Per Student Spending`, y=(ela_proficient/ela_tests)))
```

#Teacher Retention

## Current Year
Would be interesting to look at this outlier schools that have low retention but high scores.
```{r}
#Perf_Teachers_Prior_Year
ggplot(dc) + geom_point(aes(x=Perf_Teachers_Current_Year, y=(math_proficient/math_tests)))
```
## Prior Year

```{r}
#Perf_Teachers_Prior_Year

ggplot(dc) + geom_point(aes(x=Perf_Teachers_Prior_Year, y=(math_proficient/math_tests)))

```

# Test Scores by Demo
## Number of students by Demo
```{r}
dc_demo <- dc %>% select(select_vars(names(dc), ends_with("_Total")), math_proficiency, Enroll_Total) %>%
  gather(key="Demo", value="Students", select_vars(names(dc_demo), starts_with("Demo")))

ggplot(dc_demo) + geom_col(aes(x=reorder(Demo, Students), y=Students)) + coord_flip()
```

# Model Building
Variables of interest:
- Truancy
- Teacher Retention
- Demo

```{r}

fit <- lm(data=dc, math_proficiency ~ (Demo_Black))
summary(fit)

grid <- dc %>% data_grid(Demo_Black) %>% add_predictions(fit)

ggplot(dc, aes(Demo_Black, math_proficiency)) + 
  geom_point() + 
  geom_line(data = grid, aes(y = pred))

dc <- dc %>% add_residuals(fit)
ggplot(dc) + geom_point(aes(x=Demo_Black,y=resid)) 
# Is there a way to add a line based on intercept, coef alone?

#When I add Hispanic, the prediction for Demo_White/Asian turns negative??
#Compare below to fit <- lm(data=dc, math_proficiency ~ Demo_Black + Demo_Asian + Demo_White)
fit <- lm(data=dc, math_proficiency ~ Demo_Black + Demo_Asian + Demo_White + Demo_Hispanic)
summary(fit)

```
