---
title: "teacher-retention"
author: "Adam Ribaudo"
date: "November 11, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Overview
- Variable defined as "Retention of effective and highly effective teachers- Current Year"
- Teacher retention ranges from 75% to 100% with a median of 93%
- Doesn't seem to be a trend between retention and performance
- National average is apparently 85%
- Don't know why retention rates look so high when online articles say it's worse:https://wamu.org/story/18/10/03/high-salaries-high-turnover-d-c-public-schools-study-finds/
- In 2012, DCPS launched the Leadership Initiative for Teachers (LIFT), a five-stage career ladder on which teachers advance based on their annual performance rating. 

```{r}
summary(dc$Perf_Teachers_Current_Year)
```

```{r}
corrplot(dc %>% select(ela_proficiency, math_proficiency, Perf_Teachers_Current_Year, Perf_Teachers_Prior_Year) %>% na.omit() %>% cor(), is.corr=T)
```

```{r}
ggplot(dc) + geom_histogram(aes(x=Perf_Teachers_Current_Year))
```

```{r}
ggplot(dc) + geom_point(aes(y=math_proficiency, x=Perf_Teachers_Current_Year))
```

```{r}
ggplot(dc) + geom_point(aes(y=ela_proficiency, x=Perf_Teachers_Current_Year))
```