---
title: "truancy"
author: "Adam Ribaudo"
date: "November 11, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Overview

Modeling math & ELA % of proficient tests works well as long as the proficiency % metric is first transformed using sqrt(). 

- City trunancy report https://dcps.dc.gov/sites/default/files/dc/sites/dcps/publication/attachments/DCPS%20Annual%20Attendance%20Report_SY1516_FINAL.pdf 
- "By 9th grade, good attendance can predict graduation rates even better than 8th grade test scores."
- Typically defined as unexcused absence from school
-- Absence days / Total potential days
- Here it's defined as the % of students considered truant (10 or more unexcused absences)
- Far more schools with 0% proficiency than I would have thought

```{r}
summary(dc$Safe_Truancy_Current_Year)
```

```{r}
ggplot(dc) + geom_boxplot(aes(y=Safe_Truancy_Current_Year, x=as.factor(Dir_Ward)))
```

```{r}
ggplot(dc) + geom_histogram(aes(x=Safe_Truancy_Current_Year),binwidth = .05)
```

```{r}
ward_truancy <- dc %>% group_by(Dir_Ward) %>% 
  summarise(med_truancy = median(Safe_Truancy_Current_Year, na.rm=T), proficiency = sum(math_proficient,na.rm=T)/sum(math_tests, na.rm=T))

ggplot(ward_truancy) + geom_point(aes(x=med_truancy, y=proficiency))
```


## Model based on Truancy & Math
- Transforming math_proficiency helps

```{r}
ggplot(dc) + geom_point(aes(x = Safe_Truancy_Current_Year, y = sqrt(math_proficiency)))

fit <- lm(data=dc,sqrt(math_proficiency) ~ Safe_Truancy_Current_Year)

summary(fit)

dc <- dc %>% add_residuals(fit)
ggplot(dc) + geom_point(aes(x=Safe_Truancy_Current_Year, y=resid))

```

## Modeling on Truancy and ELA
```{r}
ggplot(dc) + geom_point(aes(x = Safe_Truancy_Current_Year, y = sqrt(ela_proficiency)))

fit <- lm(data=dc,sqrt(ela_proficiency) ~ Safe_Truancy_Current_Year)

summary(fit)

dc <- dc %>% add_residuals(fit)
ggplot(dc) + geom_point(aes(x=Safe_Truancy_Current_Year, y=resid))

```